\section{Languages as Proof-Cost Compression}

This non-normative profile material uses one lens from language history: proof cost. A language or framework decides which burdens move from human memory into syntax, type systems, runtimes, libraries, tooling, and conventions. In the agent era, syntax imitation is cheap; hidden ownership, hidden waiver policy, and unstated proof routes are expensive.

Structured programming, modular decomposition, and complexity metrics show the same pattern at smaller scale: good notation and boundaries reduce the amount of state a maintainer must hold in their head \cite{dijkstraGoto1968,parnasDecomposition1972,mccabeComplexity1976}. They do not eliminate repository proof cost. A language can prevent many local invalid states while still leaving cross-cutting questions unresolved: which team owns this path, which generated file is authoritative, which proof lane is required, and whether this refactor changed behavior. Jankurai moves those questions from reviewer memory into repository artifacts.

\begin{table}[t]
\caption{Language evolution as proof-cost compression.}
\label{tab:language-history}
\centering
\scriptsize
\begin{tabularx}{\columnwidth}{L{0.30\columnwidth} Y}
\toprule
\JKTableHeader
\textbf{Era or family} & \textbf{Repository-alignment interpretation} \\
\midrule
Machine/assembly & Exactness without reviewability is not enough; merge evidence needs semantic structure. \\
C/Unix & Power without structural memory safety increases audit burden; CISA/NSA guidance pushes teams toward memory-safe defaults \cite{cisaMemorySafeLanguages2025}. \\
Managed runtimes & Mature ecosystems help when boundaries and contracts are explicit. \\
Python/R/Julia & Excellent for model/data loops; risky when unbounded into durable product truth \cite{python314Release2025}. \\
JavaScript/TypeScript & Essential product surface; TypeScript adds a static loop to browser work \cite{typescript70Beta2026}. \\
Rust and safe systems & Strong reference-profile core because ownership and explicit errors reject many invalid states before runtime. \\
\bottomrule
\end{tabularx}
\end{table}

The agent-era lesson is that the language and repository must make invalid paths structurally harder and repair locality easier. Language choice compresses expression and local proof; continuous repository alignment compresses merge proof.
