\input{paper/tex/generated/public_repo_score_tables}

\section{Merge-Witness Readiness Scan}

The strongest field evidence in this edition is deliberately advisory. On May 6, 2026, Jankurai 0.8.8 scanned 30 public GitHub repositories against the current target profile. The paper uses a repo-local JSON source and digest receipt built from the run summary and star-rank table \cite{jankuraiPublicRepoRun2026}. This is not a certification result, not a defect claim, and not an empirical incident study. It is a posture scan asking whether prominent repositories expose merge-witness-ready ownership, proof routing, generated-zone receipts, context routing, proof freshness, and repair evidence.

The sample includes mature and widely used public projects such as Zed, Ruff, Deno, Meilisearch, Tauri, and ripgrep \cite{zedGithub2026,ruffGithub2026,denoGithub2026,meilisearchGithub2026,tauriGithub2026,ripgrepGithub2026}. Many of these projects have serious engineering cultures, active maintainers, and conventional CI. The surprising result is therefore not that small or immature repositories score low. It is that conventional maturity and popularity do not automatically produce the repository-alignment surfaces Jankurai expects.

The top observed public-repository score was 47, and every scanned public repository remained below the 50-point red band. The average score was 33.0. Across the run, Jankurai emitted 20,024 findings, including 19,651 hard findings, a hard-finding share of about 98.1\%. Those counts should be read as evidence gaps and repair opportunities under Jankurai policy, not as accusations against maintainers.

\PublicRepoScoreAggregateTable

The common missing surfaces were consistent with the paper's thesis. Repositories often had tests and CI, but did not expose an audit lane as a CI replacement artifact; did not bind owner maps, test maps, and generated-zone policy into a merge witness; did not leave fresh proof receipts for changed paths; and did not serialize repair queues in a form that a weaker implementation agent could consume. This matches the broader AI-assisted development evidence: DORA frames AI as an amplifier of existing delivery capability, not an automatic substitute for engineering discipline; security and secret-sprawl reports show plausible generated work can increase evidence risk; developer surveys and agent benchmarks show that nearly-correct output, environment setup, and context retrieval remain expensive failure modes \cite{doraStateAIAssistedSoftwareDevelopment2025,veracodeGenaiCodeSecurity2025,gitguardianSecretsSprawl2026,gitguardianSecretsSprawl2026Blog,stackOverflowAISurvey2025,sweAgentAci2024,setupBench2025,contextBench2026}.

\PublicRepoWeakDimensionTable

The category frequency makes the gap concrete. Vibe markers, security posture, boundary evidence, proof routing, Python containment, data access, and context routing dominate the top category summaries. These are not claims that the projects lack quality. They are claims that the repositories generally do not publish the specific merge-evidence interface required for a Jankurai witness.

The caveat matters. HLT marker counts, hard caps, and target-profile assumptions are strict policy proxies. A task marker, large file, Python placement, generated-boundary heuristic, or missing Jankurai CI artifact can be a false positive or a locally accepted tradeoff. The advisory scan is useful because it names repairable evidence gaps, not because it proves insecurity or nonconformance. Popularity, language choice, and conventional CI do not by themselves create repository alignment. The missing merge witness is exactly the standard gap Jankurai targets.

The full named table and a two-column score plot remain in Appendix~F. Keeping the main body aggregate-first avoids turning an exploratory readiness scan into a public quality leaderboard.
