\section{Reference Profile Comparison}

The comparison below is illustrative, policy-weighted, and non-normative. The standard is the merge witness and conformance artifact boundary, not a stack ranking. The scores assume durable product systems with web product surface, API/service behavior, operational security requirements, and long-running maintenance by agents and humans together. Alternative stacks can conform when they produce equivalent proof receipts and merge witnesses.

\textbf{The scoring model is not required to implement Jankurai Core and MUST NOT reject an otherwise conformant alternate profile.}

\begin{table}[t]
\caption{Reference profile score bands under the Jankurai scoring model.}
\label{tab:stack-ranking}
\centering
\small
\begin{tabularx}{\columnwidth}{Y R{0.18\columnwidth}}
\toprule
\JKTableHeader
\textbf{Profile family} & \textbf{Score band} \\
\midrule
Rust core + TypeScript/React/Vite + PostgreSQL & 91--97 \\
Go services + TypeScript/React/Vite + PostgreSQL & 86--94 \\
C\#/.NET + TypeScript/React/Vite + PostgreSQL & 85--93 \\
TypeScript product plane + Rust/Go compute cells + PostgreSQL & 82--94 \\
Kotlin/Java JVM + TypeScript/React/Vite + PostgreSQL & 82--92 \\
Rails, Python/FastAPI, or Elixir/Phoenix with strong boundaries & 72--90 \\
\bottomrule
\end{tabularx}
\end{table}

Sensitivity is concentrated in three weights: memory safety/security, ecosystem corpus strength, and runtime/concurrency. Increase ecosystem weight and Go, .NET, and JVM improve. Increase type-state and memory-safety weight and Rust separates. Increase product velocity and full-stack TypeScript rises, but it hits hard filters unless durable truth, generated contracts, and server ownership are explicit. The detailed decomposition is intentionally omitted from the main conformance line: these scores are reproducible policy assumptions, not empirical universals.

Detailed streaming comparisons do not belong in the main conformance line. The core rule is short: streaming clients belong behind adapters and generated event contracts. Kafka, Tansu, Iggy, Fluvio, NATS, Redis Streams, or equivalent systems can be profile choices when they emit replay, compatibility, security, observability, and migration evidence. The broker must not become hidden stack identity.
