\section{Related Work}

\begin{table*}[t]
\caption{Adjacent ecosystems and Jankurai's boundary.}
\label{tab:related-work}
\centering
\scriptsize
\begin{tabularx}{\textwidth}{L{0.22\textwidth} Y Y}
\toprule
\JKTableHeader
\textbf{Area} & \textbf{Representative work} & \textbf{Jankurai distinction} \\
\midrule
Programming structure and modularity & Structured programming, modular decomposition, complexity, Conway's law, software evolution, and theory building \cite{dijkstraGoto1968,parnasDecomposition1972,mccabeComplexity1976,conwayCommittees1968,lehmanEvolutionLaws1980,naurTheoryBuilding1985}. & Treats those lessons as repository-evidence obligations for agent-era merge, not only design advice. \\
Technical debt and refactoring & Technical debt, software aging, debt theory, debt mapping, debt management, and empirical refactoring practice \cite{cunninghamTechnicalDebt1992,parnasSoftwareAging1994,kruchtenTechnicalDebt2012,liTechnicalDebtMapping2015,brownManagingTechnicalDebt2010,murphyHillRefactoring2012}. & Recasts missing owner/proof/receipt evidence as evidence debt and binds repayment to merge witnesses. \\
Agent benchmarks and agents & SWE-bench, SWE-agent, OpenHands, Aider, SetupBench, ContextBench \cite{sweBench2024,sweAgentAci2024,openhands2024,aiderGithub2026,setupBench2025,contextBench2026}. & Standardizes the repository-local merge decision those agents must satisfy. \\
Supply-chain standards & SLSA, SPDX, OSV, OpenSSF Scorecard \cite{slsaSpec2026,spdxSpecifications2026,osvScanner2026,openssfScorecard2026}. & Uses evidence mechanics at the pull-request boundary: changed paths, owners, receipts, missing evidence, and witness decisions. \\
API and schema standards & OpenAPI, OpenAPI Generator, Protocol Buffers/Buf, JSON Schema \cite{openapiSpecification2026,openapiGeneratorUsage2026,protobufOverview2026,bufDocs2026,jsonSchemaSpec2026}. & Treats contracts as inputs to merge conformance rather than as the whole standard. \\
Conformance ecosystems & W3C process and Kubernetes conformance \cite{w3cProcess2026,kubernetesConformance2026}. & Applies maturity states, levels, fixtures, and submitted evidence to repository merge. \\
Agent instruction surfaces & \texttt{AGENTS.md}, Codex guidance, Claude memory, Cursor rules, Copilot instructions \cite{agentsMdSpec2026,openaiCodexAgentsMd2026,claudeCodeMemoryDocs2026,cursorRulesDocs2026,githubCopilotInstructions2026}. & Routes instruction surfaces through one conformance model so adapters cannot drift from proof lanes. \\
UX and observability evidence & Playwright, Storybook, WCAG, OpenTelemetry, JUnit, SARIF \cite{playwrightBestPractices2026,playwrightVisualComparisons2026,storybookVisualTests2026,storybookAccessibility2026,wcagTargetSizeMinimum2026,openTelemetryDocs2025,sarifSpec2023}. & Binds disconnected logs, screenshots, traces, and findings into evidence bundles and merge witnesses. \\
Repository ownership and policy & CODEOWNERS-style review, policy-as-code, reproducible build tools. & Combines ownership, proof routing, generated zones, permissions, receipts, and repair queues as one versioned repository interface. \\
\bottomrule
\end{tabularx}
\end{table*}

Programming history supplies the first cluster. Structured programming, information hiding, complexity measurement, Conway's law, software evolution, and theory building all explain why code is hard to change safely even when syntax improves. Jankurai does not compete with those ideas. It asks the repository to preserve their operational consequences as owner routes, boundaries, proof lanes, and repair receipts.

Technical-debt and refactoring work supplies the second cluster. Debt is not only messy code; it is accumulated interest on decisions whose consequences are no longer local, visible, or cheaply repairable. In AI-assisted repositories, missing evidence has the same shape: a change may be syntactically clean while still lacking proof of ownership, behavior, generated-source provenance, or waiver repayment. Jankurai's repair queue is a debt-management interface for merge evidence.

AI coding benchmarks and agents measure task completion, environment use, resource limits, context retrieval, and agent-computer interfaces. Supply-chain standards secure artifacts and dependencies. API/schema standards define contract surfaces. Conformance ecosystems show how standards become adoptable. Agent instruction surfaces are emerging as repository policy. UX, accessibility, observability, SARIF, and JUnit ecosystems provide receipts. ML operations work also warns that automation can carry hidden technical debt when boundaries, data dependencies, and feedback loops are implicit \cite{sculleyHiddenTechnicalDebt2015}.

Jankurai's contribution is the missing merge-time binding layer. It does not replace CI, CODEOWNERS-style review, SLSA/SPDX provenance, OpenAPI, Playwright, SARIF, JUnit, OpenTelemetry, agent instruction files, or technical-debt practice. It asks those systems to leave compatible evidence, then binds their outputs to changed paths, owners, proof lanes, missing evidence, repair queues, waiver state, and a closed merge-witness decision. The claim is therefore narrower and more practical than a new universal verifier: existing tools produce signals; Jankurai defines the repository-local object that decides whether those signals are sufficient to merge.
