\section{Language, Toolchain, and Release Bad-Behavior Matrix}

The language bad-behavior family turns common vibe-coding escape hatches into stable, reviewable obligations. The matrix below is intentionally practical: each row names behavior that should be hard to merge without local evidence, the detector or rule family that routes it, the receipt that makes it reviewable, and the default repair.

\begingroup
\tiny
\setlength{\tabcolsep}{2pt}
\renewcommand{\arraystretch}{1.06}
\begin{longtable}{L{0.10\textwidth} L{0.20\textwidth} L{0.17\textwidth} L{0.22\textwidth} L{0.13\textwidth}}
\caption{Language and toolchain bad-behavior matrix.}
\label{tab:language-bad-behavior-matrix}\\
\toprule
\JKTableHeader
\textbf{Surface} & \textbf{Hard reject examples} & \textbf{Rule or detector IDs} & \textbf{Required evidence} & \textbf{Repair default} \\
\midrule
\endfirsthead
\toprule
\JKTableHeader
\textbf{Surface} & \textbf{Hard reject examples} & \textbf{Rule or detector IDs} & \textbf{Required evidence} & \textbf{Repair default} \\
\midrule
\endhead
Runtime boundary & Reclassifying a runtime, service, adapter, or durable-truth boundary without deterministic proof of the new owner and proof lane. & \path{HLT-028-BOUNDARY-EVIDENCE-GAP}; owner/proof map audit & Owner map, test map, generated-zone status, contract evidence, and migration or compatibility receipt for the changed boundary. & Keep the boundary unchanged or add the missing owner/proof map and compatibility evidence. \\
Rust & Undocumented \texttt{unsafe}, public \texttt{unsafe fn} without a \texttt{\# Safety} contract, raw ownership reconstruction, unchecked indexing, \texttt{static mut}, broad lint suppression, dynamic shell execution. & \path{HLT-029-RUST-BAD-BEHAVIOR}; \path{rust.unsafe.*}; \path{rust.security.shell-c-dynamic} & Nearby \texttt{SAFETY:} proof, caller-obligation docs, focused negative tests, security proof for shell/FFI paths. & Remove the escape hatch or make the invariant local, documented, and tested. \\
SQL/PostgreSQL & Dynamic SQL strings, \texttt{DROP}/\texttt{TRUNCATE}/\texttt{DROP COLUMN} without safety evidence, \texttt{UPDATE}/\texttt{DELETE} without scope, schema truth enforced only by app code. & \path{HLT-030-SQL-BAD-BEHAVIOR}; \path{sql.dynamic-sql}; \path{sql.migration.destructive-no-proof}; \path{HLT-021} & Parameter binding, identifier allowlist, migration rollback or forward repair, row-count and lock plan, DB proof lane. & Parameterize, add constraints, split or stage the migration, and document recovery. \\
TypeScript & \texttt{@ts-nocheck}, \texttt{@ts-ignore}, \texttt{as any}, double casts, unchecked JSON or request bodies, disabled strictness, raw HTML, shell, or SQL sinks. & \path{HLT-031-TYPESCRIPT-BAD-BEHAVIOR}; \path{typescript.suppress.*}; \path{typescript.types.any-boundary}; \path{typescript.runtime.*} & Strict compiler proof, runtime decoder/schema, generated contract evidence, negative tests for malformed input. & Restore strictness, validate before narrowing, and replace dynamic sinks with typed or sanitized paths. \\
Docker & Docker socket mounts, \texttt{--privileged}, host namespaces, dangerous capabilities, unconfined profiles, secrets in layers, remote install pipes, mutable tags. & \path{HLT-032-DOCKER-BAD-BEHAVIOR}; \path{docker.compose.*}; \path{docker.install.unverified-remote}; \path{docker.image.mutable-tag} & Least-privilege profile, pinned digest or version, checksum/signature evidence, secret-free context, image/security scan. & Remove privilege, pin and verify inputs, move secrets to runtime, and narrow exposed ports. \\
Python & \texttt{eval}, \texttt{exec}, unsafe deserialization, \texttt{shell=True}, string-built DB execution, TLS verification bypass, broad exception handling, unapproved product-truth ownership. & \path{HLT-033-PYTHON-BAD-BEHAVIOR}; \path{python.exec.dynamic-code}; \path{python.deser.*}; \path{python.shell.dynamic}; \path{HLT-005} & Dated exception when needed, typed service boundary, safe loader, argv-array subprocess call, parameterized DB access, containment tests. & Replace dynamic execution, box the service, move truth to approved layers, and add boundary proof. \\
CI & \texttt{pull\_request\_target} executing PR head, \texttt{write-all}, secret printing, nonblocking security scans, mutable actions, privileged Docker on untrusted runners, credential caches. & \path{HLT-034-CI-BAD-BEHAVIOR}; \path{ci.github.*}; \path{ci.permissions.write-all}; \path{ci.security-scan.nonblocking}; \path{HLT-020} & Pinned actions, least-privilege permissions, trusted/untrusted separation, blocking security jobs, artifact provenance, workflow owner review. & Split trust contexts, minimize token scopes, pin actions, and make security failures blocking. \\
Git automation & \texttt{git reset --hard}, destructive clean, stash hidden state, force push, broad staging, \texttt{--no-verify}, destructive ref edits, credential-bearing remotes. & \path{HLT-035-GIT-BAD-BEHAVIOR}; \path{git.destructive.*}; \path{git.stage.unbounded}; \path{git.remote.*} & Status receipt, explicit path list, no hidden state, reviewed ref operation, exact staged-path evidence. & Replace broad mutation with path-scoped commands and keep verification enabled. \\
Git tools, hooks, and agents & Hooks, MCP tools, skills, or IDE/terminal agents with broad mutation authority, unknown provenance, dynamic remote instructions, or no receipts. & \path{HLT-024-AGENT-TOOL-SUPPLY-GAP}; \path{HLT-035-GIT-BAD-BEHAVIOR}; \path{HLT-012} & Tool identity, supply-chain review, least-privilege permission profile, changed-path receipt, stop conditions, and rollback plan. & Pin and review the tool, reduce permissions, route mutations through explicit receipts, or disable the hook/agent. \\
Release & Retagging, forced tag pushes, overwritten assets, skipped tests or package verification, mutable \texttt{latest}-only outputs, secret-bearing artifacts, unverified release creation, missing checksums/SBOM/provenance. & \path{HLT-037-RELEASE-BAD-BEHAVIOR}; \path{HLT-025-RELEASE-READINESS-GAP} & Version source, changelog, protected tag/commit identity, CI release receipt, checksum/SBOM/provenance/signature evidence, rollback and monitoring plan. & Publish immutable versioned artifacts from a green commit, attach integrity evidence, and document recovery. \\
\bottomrule
\end{longtable}
\endgroup
